%*******************************************************
% Abstract in English
%*******************************************************
\pdfbookmark[1]{Abstract}{Abstract}


\begin{otherlanguage}{american}
	\chapter*{Abstract}
	The \ac{IOT} gets more and more part of the everyday life. Smart-Home solutions, latency-sensitive connected cars or the vision of a smart-city affect the research works in their respective areas. The aim is a fully-automated \ac{M2M} handling of processes to automate and simplify our everday life. This leads to a huge amount of data needs to be processed, transmitted and stored. With the introduction of the new mobile communication standard 5G and increasingly powerful devices the transmission and processing of data is guaranteed. The question where this huge amount of data is stored and further processed is still unanswered.\\
	Besides the topic of \acf{DLT} gets more attention as new use cases arise which can profit from the distributed infrastructure, the trustless environment and the decentralization. It makes sense to investigate to see if the yet relativly young technologies \ac{IOT} and \ac{DLT} can complement each other and where there may be synergies.\\
	This master thesis evaluates a selection of established \acp{DLT} for their suitability for the \ac{IOT} environment with focus on the \ac{M2M} part. The key research object is the thesis, that DLT is a suitable technology for IOT and the non-functional requirements for all \ac{DLT}-\ac{IOT} use cases are identical. For this purpose an \ac{IOT} use case will be created which will be used for further analysis. Using a classification model requirements are identified that a \ac{DLT} must satisfy in order to be suitable for the use case. The criteria are applied, evaluated and ranked to a selection of \ac{DLT} implementations. As a result of the analysis, Ethereum is the most suitable solution and therefore a prototypical implementation is made to check the result of the requirements evaluation and finally the research questions.\\
	The result is a structured and comprehensible rating of multiple established \acp{DLT} and their suitability for \ac{IOT} use cases as well as a \ac{DLT} based prototype inspired by a real use case which serves as a verification for the developed rating. It is shown that \ac{IOT} can profit from \ac{DLT} under certain circumstances. The findings of this work are, that \ac{DLT} is a suitable technology for decentralized and asynchronous \ac{IOT} use cases where multiple distrusting parties are involved. The base of all non-functional requirements are identical for all DLT-\ac{IOT} usecases.
\end{otherlanguage}
